---
title: "Eruptions and Volcanoes"
author: Irene Grone
output: 
  flexdashboard::flex_dashboard:
    theme: 
      bg: "darkgrey"
      fg: "black"
      primary: "#8AD4C0"
      secondary: "lightblue"
      navbar-bg: "#FFB46E"
    orientation: rows
    vertical_layout: fill
    navbar:
      - { title: "Profile", 
          href: "http://www.linkedin.com/in/irenegrone", 
          align: right, icon: fa-linkedin }
      - { title: "Code", 
          href: "https://github.com/irenegrone", 
          align: right, icon: fa-github }
---


```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(readxl)
library(magrittr)
library(lubridate)
library(ggplot2)
library(flexdashboard)
library(leaflet)
library(plotly)


# File paths
Eruption_FILE_PATH = "data/GVP_Eruption_Results.xlsx"
Holocene_FILE_PATH = "data/GVP_Volcano_List_Holocene.xlsx"
Pleistocene_FILE_PATH = "data/GVP_Volcano_List_Pleistocene.xlsx"

# Eruptions ####
eruption_data <- read_xlsx(Eruption_FILE_PATH,
                           sheet = "Eruption List",
                           skip = 1)[, c("Volcano Number",
                                         "Volcano Name",
                                         "Eruption Number",
                                         "VEI",
                                         "Start Year",
                                         "Start Month",
                                         "Evidence Method (dating)",
                                         "Latitude",
                                         "Longitude")]

names(eruption_data) <- str_to_lower(gsub(" ", "_", names(eruption_data)))


# missing start_month are replaced with month = 1
# missing VEI are replaced with rank unknown
eruption_data <- eruption_data %>%
  rename_at("evidence_method_(dating)", ~"evidence_method_dating") %>%
  mutate(
    vei = as.numeric(vei),
    start_year = as.numeric(start_year),
    start_month = as.numeric(start_month),
    evidence_method_dating = replace_na(evidence_method_dating, "unknown"),
    # assumption missing Explosivity Index are of rank 0
    vei = replace_na(vei, "unknown"),
    # missing `Start Month` are replaced with default 1
    start_month = replace_na(start_month, 1)
  ) %>%
  filter(!is.na(start_year)) %>%
  filter(volcano_name != "Unknown Source")


# Eruption events ####
events_data <- read_xlsx(Eruption_FILE_PATH,
                                  sheet = "Events",
                                  skip = 1)[, c("Volcano Number",
                                                "Volcano Name",
                                                "Eruption Number",
                                                "Eruption Start Year",
                                                "Event Number",
                                                "Event Type")]


names(events_data) <- str_to_lower(gsub(" ", "_", names(events_data)))


events_data <- events_data %>%
  mutate(
    eruption_start_year = as.numeric(eruption_start_year)
  ) %>%
  filter(volcano_name != "Unknown Source")


# Holocene ####
holocene <- read_xlsx(Holocene_FILE_PATH,
                                sheet = "Holocene Volcano List",
                                skip = 1)


names(holocene) <- str_to_lower(gsub(" ", "_", names(holocene)))


holocene <- holocene %>%
  rename_at("elevation_(m)", ~"elevation") %>%
  select(!c(last_known_eruption, activity_evidence)) %>%
  mutate(
    dominant_rock_type = replace_na(dominant_rock_type, "Unknown"),
    tectonic_setting = replace_na(tectonic_setting, "Unknown"),
    primary_volcano_type = gsub("\\(es\\)", "", primary_volcano_type),
    primary_volcano_type = gsub("\\(s\\)", "", primary_volcano_type),
    primary_volcano_type = gsub("\\?", "", primary_volcano_type),
    primary_volcano_type = str_trim(primary_volcano_type),
  )  %>%
  select(volcano_number, volcano_name, latitude, longitude,
         region, country, primary_volcano_type, tectonic_setting,
         elevation, dominant_rock_type) %>%
  mutate(
    era = "Holocene",
  )


# Pleistocene ####
pleistocene <- read_xlsx(Pleistocene_FILE_PATH,
                         sheet = "Pleistocene Volcano List",
                         skip = 1)[ , c("Volcano Number",
                                        "Volcano Name",
                                        "Country",
                                        "Primary Volcano Type",
                                        "Region",
                                        "Subregion",
                                        "Latitude",
                                        "Longitude",
                                        "Elevation (m)",
                                        "Tectonic Setting")]


names(pleistocene) <- str_to_lower(gsub(" ", "_", names(pleistocene)))


pleistocene <- pleistocene %>%
  rename_at("elevation_(m)", ~"elevation") %>%
  mutate(
    tectonic_setting = replace_na(tectonic_setting, "Unknown"),
    primary_volcano_type = gsub("\\(es\\)", "", primary_volcano_type),
    primary_volcano_type = gsub("\\(s\\)", "", primary_volcano_type),
    primary_volcano_type = gsub("\\(\\?\\)", "", primary_volcano_type),
    primary_volcano_type = gsub("\\?", "", primary_volcano_type),
    primary_volcano_type = str_trim(primary_volcano_type),
    dominant_rock_type = "-"
  ) %>%
  select(volcano_number, volcano_name, latitude, longitude,
         region, country, primary_volcano_type, tectonic_setting,
         elevation, dominant_rock_type) %>%
  mutate(
    era = "Pleistocene",
  )



# unified volcanoes dataset
volcanoes_geo <- holocene %>%
  union(pleistocene) %>%
  separate(tectonic_setting, c("tectonic_zone", "tectonic_crust"), " / ") %>%
  mutate(
    tectonic_crust = gsub("crust ", "", tectonic_crust)
  ) %>%
  mutate(
    primary_volcano_type = as.factor(primary_volcano_type),
    dominant_rock_type = as.factor(dominant_rock_type),
    tectonic_zone = as.factor(tectonic_zone),
    tectonic_crust = as.factor(tectonic_crust),
    era = as.factor(era)
  )
```




Map {data-icon="fa-globe"}
=====================================

### <b><font size=12>Volcanoes in the World</font></b>

```{r}
leaflet(volcanoes_geo,  padding = 10) %>%
  addTiles() %>%
  setView(lng = -0, lat = 0, zoom = 1.5) %>%
  addCircleMarkers(lng = ~longitude,
                  lat = ~latitude,
                  radius = 1,
                  color = ifelse(volcanoes_geo$era == "Holocene",
                                 "#66C4AB", "royalblue"),
                  ) %>%
  addLegend("bottomright", 
            colors = c("#66C4AB", "royalblue"),
            labels = c("Holocene", "Pleistocene"),
            title = "Era"
            )
```


Volcanoes 
=====================================


Value Boxes {data-width=200}
-------------------------------------

### All Volcanoes

```{r}
cnt_volc <- length(unique(volcanoes_geo$volcano_number))

valueBox(cnt_volc, caption = "ALL VOLCANOES", color = "#FFAAA6")
```


### Holocene

```{r}
cnt_holo <- length(unique(holocene$volcano_number))

valueBox(cnt_holo, caption = "HOLOCENE VOLCANOES", color = "#8AD4C0")
```


### Pleistocene
```{r}
cnt_plei <- length(unique(pleistocene$volcano_number))

valueBox(cnt_plei, caption = "PLEISTOCENE VOLCANOES", color = "royalblue")
```


### primary_volcano_type

```{r}
cnt_type <- length(unique(volcanoes_geo$primary_volcano_type))

valueBox(cnt_type, caption = "PRIMARY VOLCANO TYPE", color = "#AD80C4")
```


### country

```{r}
cnt_country <- length(unique(volcanoes_geo$country))

valueBox(cnt_country, caption = "COUNTRIES", color = "#79BEE0")
```


Row {data-height=500}
-----------------------------------------------------------------------

### PRIMARY VOLCANO TYPE

```{r, fig.width = 6, fig.height= 5}
p <- volcanoes_geo %>%
  count(era, primary_volcano_type, .drop=FALSE) %>%
  ggplot(aes(fill = era,
             x = primary_volcano_type,
             y = n)
  ) +
  geom_bar(position="dodge",
           stat="identity"
  ) +
  scale_fill_manual("Era", values=c("#8AD4C0", "royalblue")) +
 # ylim(0, 740) +
  labs(
    x = "Volcano type",
    y = "Count",
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 0.5),
       legend.position = c(.90, .90),
        )

ggplotly(p)
```

>Number of primary volcano type by Era: the most common type is Stratovolcano.


### VOLCANOES BY REGION
```{r}
region <- volcanoes_geo %>%
  count(region, .drop=FALSE) %>%
  ggplot(aes(fill = "#FFB46E",
             x = region,
             y = n)
  ) +
  geom_text(aes(label=n),
            color = "black",
            size = 3.5
  ) +
  geom_bar(position="dodge",
           stat="identity"
  ) +
  scale_fill_manual(values=c("#FFB46E")) +
  labs(
    x = "Region",
    y = "Count",
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

ggplotly(region) %>% layout(hovermode=F)
```
>Total number of volcanoes by region.


Row {data-height=500}
-----------------------------------------------------------------------


### ELEVATION DISTRIBUTION

```{r}
elev_density <- volcanoes_geo %>%
  ggplot(aes(elevation,
             fill = era
             )
  ) +
  geom_density(alpha = 0.3) +
  scale_fill_manual("Era", values=c("#8AD4C0", "royalblue")) +
  theme_minimal()


ggplotly(elev_density)
```



>Distribution of volcano elevations by Era.


### TECTONIC SETTING ELEVATION DISTRIBUTION
```{r}
tect_col <- c("#701510", "lightgrey", "#17732F", "#94DAFF")

tect_density <- volcanoes_geo %>%
  filter(tectonic_zone != "Unknown") %>%
  ggplot(aes(elevation,
             fill = tectonic_crust,
             )
  ) +
  geom_density(alpha = 0.2) +
  facet_grid(vars(era), vars(tectonic_zone)) +
  scale_fill_manual("Crust type", values=tect_col) +
  theme_minimal()

ggplotly(tect_density)
```

>Elevation distribution in tectnonic zone split by cust type in each era.



Eruptions
=====================================

### Chart C

```{r}

```





{data-icon="fa-info-circle"}
=====================================


Row {data-height=200}
-------------------------------------

### <b><font size=12>About</font></b>

Volcanoes and eruptions are a fascinating and very interesting subject and the Smithsonian Institution (see Data section below) provide a well maintained and updated database open to the public and researchers.

This dashboard has been created as part of a data science / analytics portfolio (see github repository for code from link on top right of page). The aim of the repository is to practice visualisation with Flexdashboard on open source data.


Row 
-------------------------------------

### <b><font size=12>Data</font></b>

The data from Smithsonian Institution are available for use with the [CC0 1.0 Universal licence](https://creativecommons.org/publicdomain/zero/1.0/)
and follow the Smithsonian term of use available at https://www.si.edu/termsofuse


GVP_Volcano_List_Holocene.xlsx<br>
<cite>Global Volcanism Program, 2013. Volcanoes of the World, v. 4.9.4. Venzke, E (ed.). Smithsonian Institution. Downloaded 10 May 2021. https://doi.org/10.5479/si.GVP.VOTW4-2013 <br>
downloaded at https://volcano.si.edu/list_volcano_holocene.cfm </cite>


GVP_Volcano_List_Pleistocene<br>
<cite>Global Volcanism Program, 2013. Volcanoes of the World, v. 4.9.4. Venzke, E (ed.). Smithsonian Institution. Downloaded 10 May 2021. https://doi.org/10.5479/si.GVP.VOTW4-2013 <br>
downloaded at  https://volcano.si.edu/list_volcano_pleistocene.cfm </cite>


GVP_Eruption_Results.xlsx<br>
<cite>Global Volcanism Program, 2013. Volcanoes of the World, v. 4.9.4. Venzke, E (ed.). Smithsonian Institution. Downloaded 10 May 2021. https://doi.org/10.5479/si.GVP.VOTW4-2013 <br>
downloaded at https://volcano.si.edu/database/search_eruption_results.cfm </cite>
